Extreme heterogeneity of influenza virus infection in single cells

### 2018-01-15 paper/ResponseToReviews.tex

- Line 329: "characterizing" misspelled.

{\color{black}
This typo has been fixed.
}

- Lines 330-4: This statement about a confluence of factors leading to extreme cell-to-cell heterogeneity should be reworded since the data here actually doesn't show a clear role for innate immunity in promoting heterogeneity and there is no significant evaluation of the effect of host cell state on heterogeneity.

{\color{black} 
We have re-worded the sentences in question (the last paragraph of the Discussion).
We now make clear that we find little effect of innate-immune activation in our experiments.
We do still mention that we find \emph{at least} a minor role for host-cell state, since the new experiments described above show that activation of oxidative stress response genes does appear to play at least some causative role.
}

### 2018-01-15 paper/paper.tex
revised last paragraph of Discussion as requested by reviewers

Overall, our work shows the power and importance of characterizing cellular infection at the level of single cells~\citep{avraham2015pathogen}.
The dynamics of viral infection in any given cell is shaped by the genetic composition of the incoming virion, the host-cell state, the bi-modality of innate-immune activation, and the inherent stochasticity of molecular processes initiated by a single copy of each viral gene.
We have shown how the confluence of these factors leads to extreme cell-to-cell heterogeneity in the transcriptional outcome of influenza virus infection.
Further deconstruction of the contributions of each factor will enable a deeper understanding of how the bulk features of infection emerge from the processes occurring within individual virus-infected cells.

>>>>

Overall, our work shows the power and importance of characterizing cellular infection at the level of single cells~\citep{avraham2015pathogen}.
Viral infection can involve heterogeneity in the genetic composition of the incoming virion, the host-cell state, the bi-modality of innate-immune activation, and the inherent stochasticity of molecular processes initiated by a single copy of each viral gene.
In our experiments with short-timeframe and low-MOI infections with a relatively pure stock of influenza virus, we find only a minor role for innate-immune activation, but a substantial role for heterogeneity in the complement of viral genes that are expressed in individual cells and at least some contribution of host-cell state.
Our current experiments are not able to quantify the role of other possibly important factors such as mutations in viral genes, but we suspect that they may also contribute.
Future extensions of the approaches described here should enable further deconstruction of the sources of cell-to-cell heterogeneity during viral infection, thereby enabling a deeper understanding of how the bulk features of infection emerge from processes within individual virus-infected cells.
